---
title: "Catta"
author: "Justine Catta"
format: html
---
# link to github :
https://github.com/cattajustine-spec/test-grades.git


# introduction

```{r}
#| message: false
#| warning: false
library(vroom)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
library(readr)
library(knitr)
library(lubridate)
library(tidyr)
```

```{r}
here::i_am("test-grades.Rproj")
```

# 1.1

## question 1

```{r}
#| message: false
courses <- vroom(here("data", "courses.csv"))
```

## question 2

```{r}
courses <- courses |>
  rename(`course name` = course,
    identifier = course_id,
    `number of exams` = nb_grades) |>
  arrange(`course name`) 
 courses |> knitr::kable(format = "markdown")
```

# 1.2

## question 3

```{r}
#| message: false
students <- vroom(here("data", "students.csv"))
spec(students)
```

```{r}
students <- read_csv("data/students.csv", 
    col_types = cols(
     id = col_integer(),
    group = col_factor(),
    sex = col_factor(), 
    birth_date = col_date()))

```

```{r}
cols_to_report <- c("birth_date", "sex")

for (col in cols_to_report) {
  cat("The **", col, "** column of the **students** data frame is of class **",
      class(students[[col]])[1], "**.\n\n", sep = "")}
```

# 1.3
## question 4

```{r}
#| message: false
grades <- vroom(here("data", "grades.csv"))
```
```{r}
grades <- read_delim(
  here("data", "grades.csv"),
  delim = ";",
  na = "*",
  col_types = cols(
    id = col_integer(),
    course_id = col_integer(),
    grade = col_double()))
```

# 2
## question 5

```{r}
students |> 
  count(group) |>
  ggplot(aes(x = group, y = n)) +
  geom_bar(stat = "identity") +
  labs(x = "Group",
    y = "Number of Students",
    title = "Number of Students per Group") +
  theme_minimal()
```
It looks like there is disparities among the groups, with the group 14 having the least number of students (approx 25) and groups 4 and 5 having the most (approx 50)


## question 6

```{r}
students |>
  count(group, sex) |>
  ggplot(aes(x = group, y = n, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) +
  labs(x = "Group",
    y = "Number of Students",
    fill = "Sex",
    title = "Gender Balance per Group") +
  theme_minimal()
```
All groups have more or less the same number of girls and boys, but with some disparities. It seems like there are more boys than girls.


## question 7

```{r}
students <- students |> 
    mutate(age = round(time_length(today() - birth_date, unit = "year")))
```


```{r}
students |> 
  ggplot(aes(x = age)) +
  geom_histogram(bins = 20) +
  labs(x = "Age",
    y = "Number of Students",
    title = "Distribution of Student Ages") +
  theme_minimal()
```
A lot of students are between 21 and 23 years old, a few are around 25 and some are above 25.


## question 8

```{r}
median_age_group <- students |> 
  group_by(group) |>
  summarise("Median age" = round(median(age)), .groups = "drop")

median_age_group |> knitr::kable(format = "markdown", caption = "Median Age of Students per Group")
```

## question 9

```{r}
oldest_per_group <- students |>
  group_by(group) |>
  slice_max(age, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(id, sex, age) |>
  arrange(desc(age))

oldest_per_group |>
  knitr::kable(format = "markdown", caption = "Oldest Student in Each Group")
```
# 3
## question 10

```{r}
num_grades <- grades |>
  filter(!is.na(grade)) |>
  nrow()

cat("The data set contains", num_grades, "grades.")
```
## question 11

```{r}
courses <- courses |>
  rename(
    course = `course name`,
    course_id = identifier
  )

grades_courses <- grades |>
  inner_join(courses |> select(course_id, course),
    by = join_by(course_id)) |> select(id, course, grade) |>
  inner_join(students |> select (id, group), by = join_by(id)) |> 
  select(id, group, course, grade) 
```


```{r}
avg_grades_martial <- grades_courses |>
  filter(course == "Martial Arts and Self-Defense") |>
  group_by(group) |>
  summarise(avg_grade = mean(grade, na.rm = TRUE), .groups = "drop")
```


```{r}
avg_grades_martial |>
  ggplot(aes(x = group, y = avg_grade)) +
  geom_col() +
  labs(x = "Group",
    y = "Average Grade",
    title = "Average Grade in Martial Arts and Self-Defense per Group") +
  theme_minimal()
```

It looks like the average grade in Martial Arts is not dependent on the group, apart from group 18 that shows a smaller average than the other groups.

## question 12

```{r}
grades <- grades |>
  inner_join(courses |> select(course_id, trimester), by = join_by(course_id)) |>
  filter (!is.na(grade))

grades |>
  ggplot(aes(x = grade, fill = factor(trimester))) +
  geom_density(alpha = 0.4) +
  labs(
    x = "Grade",
    y = "Density",
    fill = "Trimester",
    title = "Distribution of Grades by Trimester"
  ) +
  theme_minimal()
```

As we can see, the peak of grades in trimester 1 is above the ones of trimester 2 and 3, because it is around 14 when the other two are more around 10 for trimester 2 and 7.5 for trimester 3. It seems like the year gets more difficult as time passes.
Also for the 1st trimester, the grades seem to be more volatile with a huge part of students between 6 and 16, the second trimester shows a small variance but with a long tail to the right.  

# 4

## question 13
```{r}
grades_per_student <- grades |>
  filter(!is.na(grade)) |>
  group_by(id) |>
  summarise(num_grades = n(), .groups = "drop") |>
  inner_join(
    students |> select(id, group, sex),
    by = join_by(id)
  )

grades_per_student |>
  slice_head(n = 5) |>
  knitr::kable(caption = "Number of grades per student (extract)")


summary_table <- grades_per_student |>
  summarise(
    min = min(num_grades),
    max = max(num_grades),
    mean = mean(num_grades),
    median = median(num_grades)
  )

summary_table |>
  knitr::kable(caption = "Summary of number of grades per student", format = "markdown")
```

## question 14

```{r}
grades_per_student |>
  ggplot(aes(x = num_grades, fill = sex)) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  labs(
    x = "Number of Grades",
    y = "Number of Students",
    fill = "Sex",
    title = "Distribution of Number of Grades by Sex"
  ) +
  theme_minimal()

grades_per_student |>
  ggplot(aes(x = num_grades, fill = sex)) +
  geom_histogram(binwidth = 0.5, position = "fill") +
  labs(
    x = "Number of Grades",
    y = "Number of Students",
    fill = "Sex",
    title = "Distribution of Number of Grades by Sex"
  ) +
  theme_minimal()
```
The number of grades is not sex-dependent as girls and boys have roughly the same amount of grades. Boys are more represented because they are more.


## question 15

```{r}
drumming_grades <- grades |>
  filter(!is.na(grade)) |>
  inner_join(
    courses |> select(course_id, course),
    by = join_by(course_id)
  ) |>
  filter(course == "Drumming and Music") |>
  group_by(id) |>
  summarise(num_grades_drumming = n(), .groups = "drop") |>
  inner_join(
    students |> select(id, group, sex),
    by = join_by(id)
  )

drumming_grades |>
  select (id, group, num_grades_drumming) |>
  rename ("Number of grades in Drumming class" = num_grades_drumming) |>
  slice_head(n = 5) |>
  knitr::kable(caption = "Number of grades in Drumming and Music per student (extract)", format = "markdown")
```
## question 16

```{r}
drumming_distribution <- drumming_grades |>
  group_by(num_grades_drumming) |>
  summarise(num_students = n(), .groups = "drop")

drumming_distribution |>
  ggplot(aes(x = num_grades_drumming, y = num_students)) +
  geom_col() +
  labs(
    x = "Number of Grades in Drumming and Music",
    y = "Number of Students",
    title = "Distribution of Number of Grades in Drumming and Music"
  ) +
  theme_minimal()
```
Most students have 9 grades, a few have 8 and a really low share of them have 7.


## question 17

```{r}
avg_grades_group <- drumming_grades |>
  group_by(group) |>
  summarise(avg_num_grades = mean(num_grades_drumming), .groups = "drop")

avg_grades_group |>
  ggplot(aes(x = group, y = avg_num_grades)) +
  geom_col() +
  labs(
    x = "Group",
    y = "Average Number of Grades",
    title = "Average Number of Grades in Drumming per Group"
  ) +
  theme_minimal()
```
As we can see, there is no link between the group and the number of grades in Drumming, per student because all have around 8,8.

## question 18

```{r}
avg_grades_group_sex <- drumming_grades |>
  group_by(group, sex) |>
  summarise(avg_num_grades = mean(num_grades_drumming), .groups = "drop")


avg_grades_group_sex |>
  ggplot(aes(x = group, y = avg_num_grades, fill = sex)) +
  geom_col(position = "dodge") +
  labs(
    x = "Group",
    y = "Average Number of Grades",
    fill = "Sex",
    title = "Average Number of Grades in Drumming by Group and Sex"
  ) +
  theme_minimal()
```
Fortunately, there seems to be no link between sex, group and grades.


## question 19

```{r}
grades_student_course <- grades_courses |>
  filter(!is.na(grade)) |>
  group_by(id, group, course) |>
  summarise(avg_grade = round(mean(grade),2), .groups = "drop") |>
  pivot_wider(
    id_cols = c(id, group),
    names_from = course,
    values_from = avg_grade
  )

grades_student_course |>
  slice_head(n = 5) |>
  knitr::kable(caption = "Average grades per student")
```

## question 20

```{r}
grades_student_course |>
  ggplot(aes(
    x = `Astronomy and Celestial Navigation`,
    y = `Storytelling and Oral Tradition`
  )) +
  geom_point(color = "steelblue", alpha = 0.7) +
  labs(
    x = "Average Grade in Astronomy and Celestial Navigation",
    y = "Average Grade in Storytelling and Oral Tradition",
    title = "Comparison of Average Grades per Student"
  ) +
  theme_minimal()
```
There is no causality between the average grade in astronomy and the average grade in storytelling. 

## question 21

```{r}
correlation_group <- grades_student_course |>
  group_by(group) |>
  summarise(
    Correlation = cor(
      `Earth Magic and Fertility Rituals`,
      `Herbalism and Traditional Medicine`
    ),
    .groups = "drop"
  )

correlation_group |>
  knitr::kable(caption = "Correlation between Earth Magic and Herbalism by group", fromat = "markdown")
```

## question 22

```{r}
grades_student_course |>
  filter(group == correlation_group |>
           slice_max(abs(Correlation), n = 1) |>
           pull(group)) |>
  ggplot(aes(
    x = `Herbalism and Traditional Medicine`,
    y = `Earth Magic and Fertility Rituals`
  )) +
  geom_point(color = "steelblue", alpha = 0.7) +
  labs(
    x = "Average Grade in Herbalism and Traditional Medicine",
    y = "Average Grade in Earth Magic and Fertility Rituals",
    title = paste("Correlation Scatter Plot for the group with the highest absolute Correlation")
  ) +
  theme_minimal()
```
We can see there is a negative correlation. 


## question 23

```{r}
final_grade <- grades_student_course |>
  rowwise() |>
  mutate(final_grade = mean(c_across(-c(id, group)), na.rm = TRUE)) |>
  ungroup() |>
  inner_join(
    students |> select(id, sex),
    by = join_by(id)
  ) |>
  select(id, group, sex, final_grade) |>
  arrange(desc(final_grade))

final_grade |>
  slice_head(n = 5) |>
  knitr::kable(caption = "Top 5 Students by Final Grade")
```
## question 24

```{r}
final_grade |>
  ggplot(aes(x = factor(group), y = final_grade)) +
  geom_boxplot(fill = "steelblue", alpha = 0.7) +
  labs(
    x = "Group",
    y = "Final Grade",
    title = "Distribution of Final Grades by Group"
  ) +
  theme_minimal()
```
The distribution of final grade by group is really volatile across groups. The highest-average group is group 12 with an average between 11.5 and 12, and the lowest-average group is 18 with an average at 10. The rest of groups is around 10.5 and 11.5.

## question 25


```{r}
final_grade |>
  ggplot(aes(x = factor(group), y = final_grade, fill = sex)) +
  geom_boxplot() +
  labs(
    x = "Group",
    y = "Final Grade",
    fill = "Sex",
    title = "Distribution of Final Grades by Group and Sex"
  ) +
  theme_minimal()
```
In some groups, girls show higher average grades and in others, boys show higher average grade, so we cannot say anything. Globally their standard deviation seems to be roughly the same for each group.

## question 26

```{r}
pass_data <- grades_student_course |>
  pivot_longer(cols = -c(id, group), names_to = "course", values_to = "avg_grade") |>
  inner_join(grades |> select(course_id, trimester) |> distinct() |> 
               inner_join(courses |> select(course, course_id), by = join_by(course_id)),
             by = join_by(course)) |>
  group_by(id) |>
  summarise(
    final_grade = mean(avg_grade, na.rm = TRUE), 
    min_course_grade = min(avg_grade, na.rm = TRUE), 
    trimester_avg_ok = all(tapply(avg_grade, trimester, mean) >= 10),
    .groups = "drop"
  ) |>
  mutate(
    pass = min_course_grade >= 5 & trimester_avg_ok
  ) |>
  inner_join(students |> select(id, group, sex), by = join_by(id)) |>
  select(id, group, sex, final_grade, pass)


pass_data |>
  slice_head(n = 5) |>
  knitr::kable(caption = "Pass/Fail Status per Student")
```

## question 27

```{r}
fail_high_grade <- pass_data |>
  filter(!pass & final_grade >= 10)

cat("The number of students who do not pass but have a final grade above 10 is", nrow(fail_high_grade))
```

## question 28

```{r}
pass_rate_group <- pass_data |>
  group_by(group) |>
  summarise(pass_rate = mean(pass), .groups = "drop")

pass_rate_group |>
  ggplot(aes(x = factor(group), y = pass_rate)) +
  geom_col() +
  labs(
    x = "Group",
    y = "Pass Rate"
  ) +
  theme_minimal()

```
The pass rate is very low, for each group. For only 7 groups, some students passed.


## question 29

```{r}
course_failures <- grades_courses |>
  filter(!is.na(grade)) |>
  filter(grade < 5) |>
  group_by(course) |>
  summarise(num_failures = n(), .groups = "drop") |>
  arrange(desc(num_failures))

course_failures |>
  slice_head(n = 5) |>
  knitr::kable(caption = "Top Courses Responsible for Failures (<5)")

```

## question 30

```{r}
age_pass <- pass_data |>
  inner_join(students |> select(id, age), by = "id") |>
  group_by(age) |>
  summarise(success_rate = mean(pass), .groups = "drop")

age_pass |>
  ggplot(aes(x = age, y = success_rate)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "red") +
  labs(
    x = "Age",
    y = "Success Rate"
  ) +
  theme_minimal()

```
It does not seem like the success rate is age dependent as very few students are older than 25 years old and most students are between 20 and 25. Here the success rate is high between 20 and 25 but not for one age in particular. 










